Kinetic Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm

### 2022-02-28 2022/PAVT/PAVT2022.tex
corrections from Ufa

\textcolor[rgb]{1,0,0}{As it has been already mentioned above, from the formal point of view we consider the inverse problem of chemical kinetics as a global optimization problem. 
In the specific problem under consideration, the values of the objective function are calculated by solving stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}). Since the right parts of the system are continuous functions with bounded derivatives, then theoretically its solution will also be continuous and bounded. Therefore, the discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.}

@@

According to the problem statements (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}), the corresponding calculations were carried out in this work.
%ННГУ
To carry out the numerical experiments, UNN supercomputer Lobachevsky was used (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node). The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.
\textcolor[rgb]{1,0,0}{The accuracy of solving the ODE system (\ref{eq:one})--(\ref{eq:twelve}) was set small enough so that the final error of the solution was much less than the accuracy of the stopping criterion of the optimization method and did not affect the method used.}

@@

\textcolor[rgb]{1,0,0}{Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) depends significantly on the estimation of the Lipschitz constant of the objective function. The constant is adaptively estimated during the work of the algorithm and may vary depending on the accumulated search information. With a correct estimate of the Lipschitz constant the method can converge to a global optimum point faster than in the case of its incorrect estimate. This explains the effect of superlinear speedup observed in experiments.}

@@

After calculating the rate constants for different temperature values, it is possible to calculate the parameters $E_i$ and $k_i^0$. So, for example, for the first stage of the reaction $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, for the third stage $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$. However, for some stages, the parameter values were fairly high $k^0$, namely for the second stage $k_2^0 = 8.52 \cdot 10^{25}$ min$^{-1}$, for the seventh stage $k_7^0 = 2.82 \cdot 10^{26}$ kg mol$^{-2}$ min$^{-2}$. Therefore, in future works, the constants corresponding to these kinetic parameters will be recalculated.

>>>>

As it has been already mentioned above, from the formal point of view we consider the inverse problem of chemical kinetics as a global optimization problem. 
In the specific problem under consideration, the values of the objective function are calculated by solving stiff ODE system (\ref{eq:one})--(\ref{eq:twelve}). Since the right parts of the system are continuous functions with bounded derivatives, then theoretically its solution will also be continuous and bounded. Therefore, the discrepancy (\ref{eq:objective_func2}) will satisfy the Lipschitz condition with a priori unknown constant.

@@

According to the problem statements (\ref{eq:objective_func1}) and (\ref{eq:objective_func2}), the corresponding calculations were carried out in this work.
%ННГУ
To carry out the numerical experiments, UNN supercomputer Lobachevsky was used (CentOS 7.2, SLURM, two CPU Intel Sandy Bridge E5-2660 2.2 GHz and 64 Gb RAM on the node). The asynchronous global optimization algorithm was implemented using C++ (GCC 5.5.0 and Intel MPI were used); the objective function values were computed using Python 3.9.
The accuracy of solving the ODE system (\ref{eq:one})--(\ref{eq:twelve}) was set small enough so that the final error of the solution was much less than the accuracy of the stopping criterion of the optimization method and did not affect the method used.

@@

Note that when solving the global optimization problem, the number of iterations of the parallel algorithm (and hence its speedup) depends significantly on the estimation of the Lipschitz constant of the objective function. The constant is adaptively estimated during the work of the algorithm and may vary depending on the accumulated search information. With a correct estimate of the Lipschitz constant the method can converge to a global optimum point faster than in the case of its incorrect estimate. This explains the effect of superlinear speedup observed in experiments.

@@

After calculating the rate constants for different temperature values, it is possible to calculate the parameters $E_i$ and $k_i^0$. So, for example, for the first stage of the reaction $E_1 = 24.65$ kJ/mol, $k_1^0 = 6.85 \cdot 10^4$ min$^{-1}$, for the third stage $E_3 = 75.57$ kJ/mol, $k_3^0 = 2.62 \cdot 10^{14}$ kg $\cdot$mol$^{-1}\cdot$min$^{-1}$. However, for some stages, the parameter values $k^0$ were fairly high, namely for the second stage $k_2^0 = 8.52 \cdot 10^{25}$ min$^{-1}$, for the seventh stage $k_7^0 = 2.82 \cdot 10^{26}$ kg mol$^{-2}$ min$^{-2}$. Therefore, in future works, the constants corresponding to these kinetic parameters will be recalculated.
Since the problem is multidimensional, there may be several solutions that describe experimental data within a certain deviation error. The solution may change if, for example, additional experimental data are found. However, the algorithm proposed in the paper provides a single solution within the search range, and high values of pre-exponential multipliers explain the presence of complex stages, which in turn consist of several stages that can be refined by chemists.
